Marriage Data analysis

### 2024-09-27 paper/paper.qmd
Cross references changed

There are also seven other detailed feature in the datasets: "ORGANIZATION_NAME", "SHELTER_NAME", "SHELTER_ADDRESS", "SHELTER_PROVINCE", "SHELTER_POSTAL_CODE", "FACILITY_NAME", and "PROGRAM_NAME". Using R programming lanaguage @citeR, the `janitor`@R-janitor, `tidyverse` @R-tidyverse, and `dplyr` @R-dplyr pakages are used in data simulation, downloading, cleaning, and writting test. No new variables were created for this analysis, but the data was aggregated by grouping observations by city, date, and sector. This aggregation facilitated a more comprehensive comparison across cities and sector groups. A random sample of the cleaned data can be seen in @fig-sample-shelter.
```{r}
#| label: fig-sample-shelter
#| tbl-cap: Sample of Cleaned Shelter Data
#| echo: false

@@

kable(shelter_data[indices,])

@@

## Data Breakdown
As shown in @fig-shelter-full-table, The four cities being observed in the dataset are Toronto, Etobicoke, North York, and Scarborough. The sectors recorded include Co-ed (mixture), Families, Men, Women, and Youth. 
```{r}
#| label: fig-shelter-full-table
#| tbl-cap: Number of Observations by City and Sector
#| echo: false
#| 
# Count the number of observations by city and sector
city_sector_breakdown <- shelter_data %>%
  group_by(shelter_city, sector) %>%
  count() %>%
  spread(key = sector, value = n, fill = 0)

@@

# Display the table with correct formatting
city_sector_breakdown %>%
  kable(caption = "Number of Observations by City and Sector, including all sectors")

@@

The following table @fig-shelter-by-city provides a summary of the average, maximum, and minimum occupancy, as well as capacity metrics for each city.
```{r}
#| label: fig-shelter-by-city
#| tbl-cap: Statistics for Homeless Shelter Occupancy by City
#| echo: false

@@

kable(combined_summary, 
             col.names = c("City", 
                           "Average Occupancy", 
                           "Max Occupancy", 
                           "Min Occupancy", 
                           "Average Capacity", 
                           "Max Capacity", 
                           "Min Capacity"))
```
The relastion ship of two key feature, occupancy and capacity, are essential for understaidng current shelter servise system. The relashipship is reflected in @fig-occupancy-capacity-relashionship. More learning on the shelter servise and thier trend, and how they reflect the fomeless population are done, and will be discuss in the next section @sec-results
```{r}
#| label: fig-occupancy-capacity-relashionship
#| fig-cap: Relationship Between Total Occupancy and Total Capacity in Shelter of the City of Toronto
#| fig-pos: "H"
#| echo: false
#| message: false

ggplot(shelter_data, aes(x = total_capacity, y = total_occupancy, color = sector)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship Between Total Occupancy and Total Capacity in Toronto Shelters",
       x = "Total Capacity",
       y = "Total Occupancy",
       color = "Sector") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

@@

## Shelter Occupancy Trend over Time
The figure
```{r}
#| label: shelter-over-time
#| fig-cap: Visualization of Shelter _____ During Time Period 2017 to 2020
#| echo: false
#| message: false

@@

# Present 4 plot together
combined_plots <- grid.arrange(toronto_plot, north_york_plot,
                                scarborough_plot, etobicoke_plot,
                                nrow = 2, ncol = 2)

combined_plots

@@

## Shelter Ustillization Analysis
To analyze capacity utilization for the four cities—Toronto, North York, Scarborough, and Etobicoke— calculate the capacity utilization rate for each city over time and visualize this data. 
```{r}
#| label: shelter-utilization
#| fig-cap: Change in Utilization Rate Average on All Shelters by City (2017-2020)
#| echo: false
#| message: false

@@

ggplot(annual_utilization, aes(x = year, y = avg_utilization_rate, color = shelter_city)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(x = "Year",
       y = "Average Utilization Rate",
       color = "City") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 100))  
```

## Over Capacity Cases
Back to the original dataset, identify cases where shelter capacity is exceeded (i.e., occupancy > capacity) and understand the frequency and distribution of these events.
```{r}
#| label: over-capacity
#| tbl-cap: Overcrowded Cases by City and Sector
#| echo: false
#| message: false

@@

# View the frequency table with knitr
frequency_table %>%
  kable()

>>>>

There are also seven other detailed feature in the datasets: "ORGANIZATION_NAME", "SHELTER_NAME", "SHELTER_ADDRESS", "SHELTER_PROVINCE", "SHELTER_POSTAL_CODE", "FACILITY_NAME", and "PROGRAM_NAME". Using R programming lanaguage @citeR, the `janitor`@R-janitor, `tidyverse` @R-tidyverse, and `dplyr` @R-dplyr pakages are used in data simulation, downloading, cleaning, and writting test. No new variables were created for this analysis, but the data was aggregated by grouping observations by city, date, and sector. This aggregation facilitated a more comprehensive comparison across cities and sector groups. A random sample of the cleaned data can be seen in @fig-shelter-sample.
```{r}
#| label: fig-shelter-sample
#| tbl-cap: Sample of Cleaned Shelter Data
#| echo: false
#| message: false

@@

kable(shelter_data[indices,],
      caption = "Sample of Cleaned Shelter Data")

@@

## Data Breakdown
As shown in @fig-shelter-full-table, The four cities being observed in the dataset are Toronto, Etobicoke, North York, and Scarborough. The sectors recorded include Co-ed (mixture), Families, Men, Women, and Youth. 
```{r}
#| label: fig-shelter-full-table
#| tbl-cap: Number of Observations by City and Sector
#| echo: false
#| message: false

# Count the number of observations by city and sector
city_sector_breakdown <- shelter_data %>%
  group_by(shelter_city, sector) %>%
  count() %>%
  spread(key = sector, value = n, fill = 0)

@@

# Display the table with correct formatting
city_sector_breakdown %>%
  kable(caption = "Number of Observations by City and Sector")

@@

The following table @fig-shelter-diff-city provides a summary of the average, maximum, and minimum occupancy, as well as capacity metrics for each city.
```{r}
#| label: fig-shelter-diff-city
#| tbl-cap: Statistics for Homeless Shelter Occupancy by City
#| echo: false
#| message: false

@@

kable(combined_summary, 
             col.names = c("City", 
                           "Average Occupancy", 
                           "Max Occupancy", 
                           "Min Occupancy", 
                           "Average Capacity", 
                           "Max Capacity", 
                           "Min Capacity"),
      caption = "Statistics for Homeless Shelter Occupancy by City")
```
The relastion ship of two key feature, occupancy and capacity, are essential for understaidng current shelter servise system. The relashipship is reflected in @fig-occupancy-capacity-relashionship. More learning on the shelter servise and thier trend, and how they reflect the fomeless population are done, and will be discuss in the next section @sec-results
```{r}
#| label: fig-occupancy-capacity-relashionship
#| fig-cap: Relationship Between Total Occupancy and Total Capacity in Shelter of the City of Toronto
#| fig-pos: "H"
#| echo: false
#| message: false

ggplot(shelter_data, aes(x = total_capacity, y = total_occupancy, color = sector)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship Between Total Occupancy and Total Capacity in Shelter of the City of Toronto",
       x = "Total Capacity",
       y = "Total Occupancy",
       color = "Sector") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

@@

## Shelter Occupancy Trend over Time
The figure @fig-shelter-over-time, a
```{r}
#| label: fig-shelter-over-time
#| fig-cap: Visualization of Shelter Occupancy During Time Period 2017 to 2020
#| echo: false
#| message: false

@@

combined_plot <- grid.arrange(toronto_plot, north_york_plot, 
                               scarborough_plot, etobicoke_plot, 
                               nrow = 2, ncol = 2,
                               top = "Visualization of Shelter Occupancy During Time Period 2017 to 2020"
                          )

# This will suppress the grid output in the document
invisible(combined_plot)

@@

## Shelter Ustillization Analysis
To analyze capacity utilization for the four cities—Toronto, North York, Scarborough, and Etobicoke— calculate the capacity utilization rate for each city over time and visualize this data. See @fig-shelter-utilization.
```{r}
#| label: fig-shelter-utilization
#| fig-cap: Change in Utilization Rate Average on All Shelters by City (2017-2020)
#| echo: false
#| message: false
#| warning: false

@@

ggplot(annual_utilization, aes(x = year, y = avg_utilization_rate, color = shelter_city)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(x = "Year",
       y = "Average Utilization Rate",
       color = "City",
       caption = "Utilization Rate Average Change on All Shelters by City") +
  theme_minimal() +
  scale_y_continuous(labels = scales::percent_format(scale = 100))  
```

## Over Capacity Cases
Back to the original dataset, identify cases where shelter capacity is exceeded (i.e., occupancy > capacity) and understand the frequency and distribution of these events. see @fig-overcapacity, under
```{r}
#| label: fig-overcapacity
#| tbl-cap: Overcrowded Cases by City and Sector
#| echo: false
#| message: false

@@

# View the frequency table with knitr
frequency_table %>%
  kable(caption = "Overcrowded Cases by City and Sector")
